Reinforcement \change{L}earning for \change{O}ptions\\on \change{T}arget \change{V}olatility \change{F}unds\thanks{Supported by ISCRA from Cineca, \email{iscra@cineca.it}.}

### 2023-02-25 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
Add introduction and abstract corrections

\title{Reinforcement \change{L}earning for \change{O}ptions\\on \change{T}arget \change{V}olatility \change{F}unds\thanks{\change{Our sincere thanks go to Marco Bianchetti and Diego Pierluigi Giovannini of Intesa Sanpaolo Milan, as well as to the Italian computing center Cineca, who generously supported our numerical simulations through an ISCRA C grant. The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.}}}

@@

\begin{abstract}
We deal with the funding costs rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio of risky assets and a risk-free one dynamically rebalanced in order to keep the realized volatility of the portfolio on a certain level. The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs \change{of its} \remove{for each} component requires to solve a control problem to evaluate the option prices. We derive an analytical solution of the problem in the Black and Scholes (BS) scenario. Then we use Reinforcement Learning (RL) techniques to determine the fund composition leading to the most conservative price under the local volatility (LV) model, for which an {\it a priori} solution is not available. We show how the performances of the RL agents are compatible with those obtained by applying path-wise the BS analytical strategy to the TVS dynamics, which therefore appears competitive also in the LV scenario.

@@

\section{Introduction}
In the recent years portfolio managers are exposed to very low interest rates and quickly changing market volatilities. An effective solution to control risks under such an environment is given by target volatility strategies (TVSs)\change{,} \remove{(}also known as constant volatility targeting\remove{)}\change{,} which are able to preserve the portfolio at a predetermined level of volatility. A TVS is \change{a strategy for managing} a portfolio of risky assets (typically  equities) and a risk-free \change{one} \remove{asset} dynamically re-balanced with the aim of maintaining the overall portfolio volatility level closed to some target value. The constant volatility approach can help investors to obtain desired risk exposures over the short and long term and \change{possibly} it increases the risk-adjusted performance of the portfolio.

These products were initially offered in the Asian markets, see for instance \remove{the reports of} \cite{Chew2011} and \cite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison2013}. At the present day we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.

In the recent literature\change{,} \remove{TVSs} \change{TVS-based portfolios} are tested to investigate their performances in term of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}. Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners. Those derivative instruments are known as target volatility options (TVOs) and have gained great interest in the derivative pricing literature, see in particular \cite{Albeverio2019}, \cite{DiGraziano2012}, \cite{DiPersio2019}, and \cite{Grasselli2016}.

This paper is the first attempt, to the best of our knowledge, to study the funding costs coming from hedging the risky assets underlying the target volatility strategy. In particular, we consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS. The portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS. Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario, i.e. the most expensive strategy in term of the financing costs. Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition. In our work we provide to the reader a formal description of the stochastic control problem and most importantly we derive an analytical solution assuming that the risky asset dynamics underlying the TVS  follow a Black and Scholes (BS) model \cite{BlackScholes1973} and that the derivative contract is a European-style option.

Although from a theoretical point of view this solution represents a sub-optimal strategy when dealing with generic dynamics for the risky assets, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem. More precisely we numerically study the problem in the general case of a local volatility dynamics \cite{Derman1994}, \cite{Dupire1994}  by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} \remove{vanilla} direct policy \change{search} algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}. 
\\

The paper is organized as follows. In Section \ref{sec:TVS} we introduce more in detail the manager-bank contract and provide the description of the TVS dynamics in presence of valuation adjustments such as the hedging costs. Then, in Section \ref{sec:Derivative} we introduce the structured class of derivative contracts linked to TVSs, where we describe the arising dynamical control problem for pricing those options. Moreover in this section we derive the BS closed solution for European TVOs in two different ways: one applying the Gy\"ongy Lemma and the other \change{through} \remove{by writing} the Hamilton-Jacobi-Bellman equation. In Section \ref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a description of the algorithm we have built. We conclude the paper with Section \ref{sec:Numerical_investigations} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility one.
\\

@@

\section{Target \change{V}olatility \change{S}trategy}\label{sec:TVS}
As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging the risky assets underlying a target volatility portfolio. Thus we consider the following scenario: a bank selling a protection to a portfolio manager who has his capital invested \change{according to} \remove{in} a TVS. In our case the fund manager has the freedom of changing the relative weights of the risky asset during the life of the TVS; once the allocation strategy is selected then the volatility targeting algorithm rebalances the risky component of the portfolio with the risk-free one in order to keep the overall portfolio volatility close to a target value. Clients investing in the fund pay a running fee for the service of the fund manager and their capital is protected.

The fund manager usually buys from a bank an option on the TVS to ensure capital protection. For instance, the capital can be protected by buying a put option. In this case, we can write the \change{undiscounted} net asset value (NAV) $A_t$ of the strategy as given by\footnote{\remove{Here we neglect discounting factors.}}
\begin{equation}
    A_t \coloneqq \max\{V_t,K\} = V_t + (K-V_t)^+ \; ,
\end{equation}
where $V_t$ is the price process of the strategy, and $K$ is the guaranteed capital. On the other hand, the fund manager can replicate the payoff by means of the put-call parity by investing the capital in a low-risk asset and buying a call on the strategy
\begin{equation}
    A_t = K + (V_t-K)^+ \; .
\end{equation}

In this way, the TVS is only defined in the two contracts client-fund and fund-bank. The fund manager is not implementing the strategy by trading in the market, and he is not subject to additional costs to access the market. On the other way, the bank is paying such costs since \change{it} \remove{she} is actively hedging the call option sold to the manager.  The bank trading activity implemented to actively hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
We remark that the choices of the manager trading activity are stochastic processes since they will depend on the market evolution. Thus, neither the manager strategy nor the bank financing costs for hedging the option can be written in the fund-bank contract. For this reason, the price of a financial product sold by the bank is adjusted to include any valuation adjustment due to the trading activity.

Our aim is to find the most expensive investing strategy from the point of view of financing costs for the bank that the manager could choose in the market. In other words, we want to determine the worst-case scenario for the bank. In the current section, we proceed by defining the price process of the TVS so that we can highlight the impact of valuation adjustments.

@@

\subsection{The Volatility Targeting Constraint}
In a typical TVS, the fund manager selects a risky-asset portfolio with a specific time-dependent allocation strategy expressed by means of the vector of relative weights $\alpha_t$, along with a risk-free asset, which we can identify with the bank account $B_t$. In the following, for sake of simplicity in exposition, we consider only total-return securities, namely we set $\pi_t=0$ \change{since no dividends are paid}. Thus we can write Equation~\eqref{eq:TVS_first} as given by
\begin{equation}
    \frac{dI_t}{I_t}= \omega_t \alpha_t \cdot \frac{dS_t}{S_t} + \left(1-\omega_t  \alpha_t \cdot \mathds{1} \right)\frac{dB_t}{B_t}- \phi_t dt \; ,
\label{eq:TVS_elegant}\end{equation}
where $\mathds{1}$ is a $n$-dimensional vector of ones and $\omega_t\in[0,1]$ is determined so that the strategy log-normal volatility is kept constant, namely
\begin{equation}
        \omega_t: \quad \mathrm{Var}_t[dI_t] = \bar{\sigma}^2I_t^2dt \; ,
\end{equation}
where $\Bar{\sigma}$ is the target volatility value.
In practice, this means that the fund manager will select a risky-portfolio choosing $\alpha_t$ equities from the universe where \change{s}he can trade and after that, \change{her} \remove{his} choices will be scaled by the automatic target volatility algorithm $\omega_t$.\footnote{We recall that the universe of assets where the manager can trade and the value of $\bar{\sigma}$ are written in the contract.}

>>>>

\title{Reinforcement \change{L}earning for \change{O}ptions\\on \change{T}arget \change{V}olatility \change{F}unds\thanks{\change{Our sincere thanks go to Marco Bianchetti and Diego Pierluigi Giovannini of Intesa Sanpaolo Milan, as well as to the Italian computing center Cineca for its support on our numerical simulations. The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.}}}

@@

\begin{abstract}
We deal with \remove{the funding costs} \change{a stochastic optimal control problem} rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio \remove{of risky assets and a risk-free one dynamically rebalanced in order to keep the} \change{whose asset-allocation is adjusted to mantain the} realized volatility of the portfolio on a certain level. \change{We consider the point of view of a derivative writer selling an option contract as protection to a portfolio manager on the capital invested in a TVS.} The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs \change{of its} \remove{for each} component requires to \remove{solve a control problem to evaluate the option} \change{adjust the protection} price\remove{s} \change{to include these costs in the worst-case scenario for the seller}. We derive an analytical solution of the problem in the Black and Scholes (BS) scenario. Then we use Reinforcement Learning (RL) techniques to determine the fund composition leading to the \remove{most conservative price} \change{optimal policy} under the local volatility (LV) model, for which an {\it a priori} solution is not available. We show how the performances of the RL agents are compatible with those obtained by applying path-wise the BS analytical strategy to the TVS dynamics, which therefore appears competitive also in the LV scenario.

@@

\section{Introduction}
In the recent years portfolio managers \remove{are} \change{were} exposed to very low interest rates and quickly changing market volatilities. An effective solution to control risks under such an environment is given by target volatility strategies (TVSs)\change{,} \remove{(}also known as constant volatility targeting\remove{)}\change{,} which are able to preserve the portfolio at a predetermined level of volatility. A TVS is \change{a strategy for managing} a portfolio of risky assets (typically  equities) and a risk-free \change{one} \remove{asset} dynamically re-balanced with the aim of maintaining the overall portfolio volatility level closed to some target value. \change{The strategy's concept involves adjusting the allocation of high-risk assets to low-risk assets during periods of high volatility in the high-risk assets, and vice versa. Thus the dynamic allocation is based on both the actual volatility of the high-risk asset and a predetermined target level of volatility. }The constant volatility approach can help investors to obtain desired risk exposures over the short and long term and \change{possibly} it increases the risk-adjusted performance of the portfolio.

These products were initially offered in the Asian markets, see for instance \remove{the reports of} \cite{Chew2011} and \cite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison2013}. At the present day we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.
In the recent literature\change{,} \remove{TVSs} \change{TVS-based portfolios} are tested to investigate their performances in term of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}\change{, where the authors propose the use of artificial neural networks for volatility forecasting volatility to enhance the performance of an asset allocation strategy}.

\remove{Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners.} \change{The success of TVS-based portfolios among investors has led to the emergence of new structured products with such portfolios as underlyings}. Those derivative instruments are known as target volatility options (TVOs) \change{and have become increasingly important in the derivative market.} \change{For these reasons TVOs}  have gained great interest in the derivative pricing literature \change{and research}, see in particular \cite{Albeverio2019}, \cite{DiGraziano2012}, \cite{DiPersio2019}, and \cite{Grasselli2016}.

This paper \change{aims to enrich the pricing framework of the TVOs by studying for the first time,} \remove{is the first attempt} to the best of our knowledge, \remove{to study} the funding costs coming from hedging the risky assets underlying the target volatility strategy. In particular, we consider the point of view of \remove{a bank} \change{an option writer (for example a bank)} selling a call option to a portfolio manager as protection on the capital invested in a TVS. \change{During her activity the} \remove{The} portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS \change{depending on how the market evolves. This mechanism is no regulated by the contract, as the manager has decision making power on which asset invest}. Since \change{the fund manager's strategy cannot be known {\it a priori} and since} the risky assets have different hedging costs, the \remove{bank} \change{option writer} shall adjust the price of the protection to include \remove{them} \change{these costs} in the worst-case scenario, i.e. the most expensive strategy in term of the financing costs. Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition\change{: determine the manager's optimal strategy that leads to the maximum protection price}.

In our work we provide to the reader a formal description of the stochastic control problem \change{and we show that it can be solved only numerically considering a general state-dependent dynamics for the risky assets.} \change{However, we prove that} \remove{and most importantly we derive} an analytical solution to the problem \change{exists,} assuming that the risky asset dynamics underlying the TVS  follow a Black and Scholes (BS) model \cite{BlackScholes1973} and that the derivative contract is a European-style option. \change{As our first contribution, we derive the BS closed solution into two different ways: applying the Gy\"ongy Lemma and the other through the Hamilton-Jacobi-Bellman equation.}

\change{In the second part of our work we tackle the problem in the general case of a state-dependent dynamics for the risky assets for which only the numerical approach is admissible.}
\remove{Although from a theoretical point of view this solution represents a sub-optimal strategy when dealing with generic dynamics for the risky assets, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem}. More precisely we numerically study the problem \remove{in the general case of  a local volatility dynamics \cite{Derman1994}, \cite{Dupire1994} by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} vanilla direct policy algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}} \change{assuming a local volatility (LV) dynamics \cite{Derman1994}, \cite{Dupire1994} \change{for the risky assets}. The reason for selecting the LV model is its widespread adoption in the industry to value derivatives, especially those related to asset baskets \cite{Gatheral2006}. The LV investigation is based on a Reinforcement Learning approach, in particular through an \textit{ad hoc} direct policy search algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}. From our study, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem under the LV dynamics.}
\\

The paper is organized as follows. In Section \ref{sec:TVS} we introduce more in detail the \remove{manager-bank} contract \change{between the option-writer and the manager} and provide the description of the TVS dynamics in presence of valuation adjustments such as the hedging costs. Then, in Section \ref{sec:Derivative} we introduce the structured class of derivative contracts linked to TVSs, where we describe the arising dynamical control problem for pricing those options. Moreover in this section we derive the BS closed solution for European TVOs in two different ways: one applying the Gy\"ongy Lemma and the other \change{through} \remove{by writing} the Hamilton-Jacobi-Bellman equation. In Section \ref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a description of the algorithm we have built. We conclude the paper with Section \ref{sec:Numerical_investigations} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility one.
\\

@@

\section{Target \change{V}olatility \change{S}trategy}\label{sec:TVS}
As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging the risky assets underlying a target volatility portfolio. Thus we consider the following scenario: a \remove{bank} \change{option writer} selling a protection to a portfolio manager who has \remove{his} \change{her} capital invested \change{according to} \remove{in} a TVS. In our case the fund manager has the freedom of changing the relative weights of the risky asset during the life of the TVS; once the allocation strategy is selected then the volatility targeting algorithm rebalances the risky component of the portfolio with the risk-free one in order to keep the overall portfolio volatility close to a target value. Clients investing in the fund pay a running fee for the service of the fund manager and their capital is protected.

The fund manager usually buys from a \remove{bank} \change{counterpart} an option on the TVS to ensure capital protection. For instance, the capital can be protected by buying a put option. In this case, we can write the \change{undiscounted} net asset value (NAV) $A_t$ of the strategy as given by\footnote{Here we neglect discounting factors \change{for sake of clarity}.}
\begin{equation}
    A_t \coloneqq \max\{V_t,K\} = V_t + (K-V_t)^+ \; ,
\end{equation}
where $V_t$ is the price process of the strategy, and $K$ is the guaranteed capital. On the other hand, the fund manager can replicate the payoff by means of the put-call parity by investing the capital in a low-risk asset and buying a call on the strategy
\begin{equation}
    A_t = K + (V_t-K)^+ \; .
\end{equation}

In this way, the TVS is only defined in the two contracts client-fund and fund-\change{option writer} \remove{fund-bank}. The fund manager is not implementing the strategy by trading in the market, and \remove{he} \change{she} is not subject to additional costs to access the market. On the other way, the \remove{bank} \change{writer} is paying such costs since she is actively hedging the call option sold to the manager.  The \remove{bank} \change{writer} trading activity implemented to actively hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
We remark that\change{, upon entering into the contract,}  the choices of the manager trading activity \remove{are} \change{can be seen as} stochastic processes since they \change{are not known {\it a priori} and }will depend on the market evolution. Thus, neither the manager strategy nor the \remove{bank} \change{writer} financing costs for hedging the option can be written in the fund-\remove{bank}\change{writer} contract\change{: the only contractual elements are the protection strike and maturity, the trading market of the manager and the target volatility of the TVS portfolio}. For this reason, the price of a financial product sold by the \remove{bank} \change{writer} is adjusted to include any valuation adjustment due to the trading activity.

Our aim is to find the most expensive investing strategy from the point of view of financing costs for the \remove{bank} \change{writer} that the manager could choose in the market. In other words, we want to determine the worst-case scenario for the bank. In the current section, we proceed by defining the price process of the TVS so that we can highlight the impact of valuation adjustments.

@@

\subsection{The Volatility Targeting Constraint}
In a typical TVS, the fund manager selects a risky-asset portfolio with a specific time-dependent allocation strategy expressed by means of the vector of relative weights $\alpha_t$, along with a risk-free asset, which we can identify with the bank account $B_t$. In the following, for sake of simplicity in exposition, we consider only total-return securities, namely we set $\pi_t=0$ \change{, which means that the fees received by the stock lender are null since no dividends are paid}. Thus we can write Equation~\eqref{eq:TVS_first} as given by
\begin{equation}
    \frac{dI_t}{I_t}= \omega_t \alpha_t \cdot \frac{dS_t}{S_t} + \left(1-\omega_t  \alpha_t \cdot \mathds{1} \right)\frac{dB_t}{B_t}- \phi_t dt \; ,
\label{eq:TVS_elegant}\end{equation}
where $\mathds{1}$ is a $n$-dimensional vector of ones and $\omega_t\in[0,1]$ is determined so that the strategy log-normal volatility is kept constant, namely
\begin{equation}
        \omega_t: \quad \mathrm{Var}_t[dI_t] = \bar{\sigma}^2I_t^2dt \; ,
\end{equation}
where $\Bar{\sigma}$ is the target volatility value.
In practice, this means that the fund manager will select a risky-portfolio choosing $\alpha_t$ equities from the universe where \change{s}he can trade and after that, \change{her} \remove{his} choices will be scaled by the automatic target volatility algorithm $\omega_t$.\footnote{We recall that the universe of assets where the manager can trade and the value of $\bar{\sigma}$ are written in the contract.}
